Dark Control:\\ A Unified Account of Default Mode Function\\ by Control Theory and Reinforcement Learning

### 2017-04-07 paper/paper.tex
fix: PC hypothesis 2.0

\subsection{Predictive coding hypothesis}
The predictive coding account
\citep{clark2013whatever, friston2008hierarchical}
is a frequently evoked idea related to the default mode function
\citep{bar2007}.
According to this cognitive account,
cortical responses emerge from continuous functional interaction between
higher and lower levels of the neural processing hierarchy.
This dynamic intercourse enables inference on the world by reconciling
gaps between fresh sensory input and stored prior information.
Feed-forward sensory experience is constantly calibrated by
top-down modulation at various hierarchical processing levels.
In fact, axonal
back projections probably outnumber by far the input projection
existing in the brain \citep{salin1995corticocortical}.
This can explain the constructive, generative nature of sensory perception
\citep{friston2010free} and
why motor action is intimately related to sensory expectations
\citep{wolpert1995internal, kording2004bayesian}.
At each stage of neural processing,
an internally generated prediction of the sensory input is
directly compared against the actual input from the environment.
A prediction error at one of the processing levels,
i.e. the difference between sensory input
and internally predicted sensation,
incurs plasticity changes of neuronal back projections (i.e., model parameters)
for gradually improved future prediction of the environment.
Contextual integration is hence maintained by top-down modulation of sensorimotor
processing by a-priori information about the environment.
The generative model of how perceived sensory events arise in the
world would be incorporated into
the current neuronal wiring.
Indeed,
This process permits updates of the internal model of the environment
to best fit the constant influx of sensory examples.

There a number of common properties between the predictive coding account
and the proposed formal account.
Both frameworks are plausibilized by neurobiological evidence that
the human brain may be viewed as a ``statistical organ'' aiming to
generalize from the past to new experiences.
Both accounts provide a parsimonious explanation why the
human brain decreases processing load devoted to incoming information
when the environment is predictable and allocate increased
computation resources when novel stimuli or unexpected events are encountered.
The predictive coding and formal account
naturally evoke a mechanisms of brain plasticity in that
neuronal wiring gets increasingly adapted
when faced by unprecedented environmental
challenges.

While sensory experience is a generative process from both views,
in predictive coding sensory perception of the external world
is a generative experience due to modulatory top down-experience at
various hierarchical levels of sensory input processing. Instead,
the Markov property of our sequential decision making framework
In our RL view, the environmental model is
\textbf{effectively expressed in the current state of the human agent
XXX guys, please double-check}.
The hierarchical processing aspect from predictive coding
is re-expressed in our account in form of
nested prediction of probable upcoming actions, states, and rewards.
Both model the consequences of action.
In the present account, the horizon of that
look into the future is explicitly expressed in the $\gamma$ parameter
in the Bellman equation.
The predictive coding framework is typically explained without
explicitly parameterization of the agent's horizon and
has a tendency to be presented as emphasizing prediction about the
immediate future.
% \item \suggestremove{Mismatch negativity of PC is immanent in reward prediction error
% by the difference between actually predicted reward of an action given
% a state and the reward predicted by non-linear regression with graded
% discounting of future rewards.}
The process of adapting the neuronal connections
for improved top-down modulation
takes the concrete form of gradient computation
back-propagation in our RL implementation,
although the neurobiological plausibility of
the back-propagation procedure is controversal
\citep{goodfellow2016deep}.
\item In sum,
the formal account may serve as a concrete implementation of
the predictive coding account
from cognitive neuroscience.
MPDs have the advantage of exposing an explicit
mechanism of the horizon of future considerations or
how the internal model of the world gets updated,
and collapses sensory input processing and action output preparation.
%
As such, recasting DMN function in terms of MDPs naturally incorporates
a majority of aspects of the prediction coding hypothesis.

>>>>

\subsection{Predictive coding hypothesis}
Predictive coding mechanisms
\citep{clark2013whatever, friston2008hierarchical}
are a frequently evoked idea in the context of default mode function
\citep{bar2007}.
According to this view,
cortical responses emerge from continuous functional interaction between
higher and lower levels of the neural processing hierarchy.
Feed-forward sensory experience is constantly calibrated by
top-down modulation realized by more multi-modal and associative brain regions
further away from the primary sensory cortices.
The dynamic interplay between cortical processing levels
may enable inferences about the world by reconciling
gaps between fresh sensory input and expectations computed
from stored prior information.
At each stage of neural processing,
an internally generated prediction of aspects of sensory input is
directly compared against the actual environmental input.
A prediction error at one of the processing levels,
i.e., the difference between sensory input
and internally predicted sensation,
incurs plasticity changes of neuronal back projections
(i.e., adapting model parameters)
to allow for gradually improved future prediction of the environment.
In fact, axonal
back projections probably outnumber by far the axonal input projections
existing in the primate brain \citep{salin1995corticocortical}.
This can explain the constructive, generative nature of sensory perception
\citep{friston2010free} and
why motor action is probably intimately related to sensory expectations
\citep{wolpert1995internal, kording2004bayesian}.
Contextual sensorimotor integration of perception-action cycles
may hence be maintained by top-down modulation
by a-priori information about the environment.

This process permits updates of the internal representation of the environment
to best accommodate and prepare for the constant influx of sensory examples.
There are hence a number of common properties
between the predictive coding account
and the proposed formal account.
Importantly,
the generative model of how perceived sensory cues arise in the
world would be incorporated into
the current neuronal wiring.
Further,
both frameworks are plausibilized by neurobiological evidence that suggest
the human brain to be a ``statistical organ'' with the biological purpose to
generalize from the past to new experiences.
Both accounts also provide a parsimonious explanation why the
human brain decreases processing load devoted to incoming information
when the environment is predictable and allocates increased
computation resources when novel stimuli or unexpected events are encountered.
The predictive coding and formal account
naturally evoke a mechanisms of brain plasticity in that
neuronal wiring gets increasingly adapted
when faced by unprecedented environmental
challenges.

While sensory experience is a generative process from both views,
the predictive coding account frames
sensory perception of the external world
as a generative experience due to the modulatory top-down influence at
various hierarchical levels of sensory input processing. Instead,
the Markov property of our sequential decision making framework.
In our RL view, the environmental model is
\textbf{effectively expressed in the current state of the human agent
XXX guys, please double-check}.
The hierarchical processing aspect from predictive coding
is re-expressed in our account in form of
nested prediction of probable upcoming actions, states, and rewards.
Both frameworks model the consequences of action.
The predictive coding framework is typically explained without
explicitly parameterization of the agent's horizon and
has a tendency to be presented as emphasizing prediction about the
immediate future.
In the present account, the horizon of that
look into the future is explicitly expressed in the $\gamma$ parameter
of the Bellman equation.
% \item \suggestremove{Mismatch negativity of PC is immanent in reward prediction error
% by the difference between actually predicted reward of an action given
% a state and the reward predicted by non-linear regression with graded
% discounting of future rewards.}
The process of adapting the neuronal connections
for improved top-down modulation
takes the concrete form of gradient computation and
back-propagation in our RL implementation.
It is however important to note that
the neurobiological plausibility of
the back-propagation procedure is controversal
\citep{goodfellow2016deep}.
\item In sum,
the formal account may serve as a concrete implementation of
the predictive coding account
from cognitive neuroscience.
MPDs have the advantage of exposing an explicit
mechanism of the horizon of future considerations or
how the internal model of the world gets updated,
and
more explicitly reconciles sensory input processing and action output preparation.
%
As such, recasting DMN function in terms of MDPs naturally incorporates
a majority of aspects of the prediction coding hypothesis.
